---
title: "DAG"
format: html
editor: visual
---

```{r}
#| include: false

library("ggdag")
library("ggplot2")
library(dagitty)
```

# Long-term HRV and CVD

```{r}
#| echo: false


HRV_cycle_DAG <- dagitty('dag {
bb="0,0,1,1"
Age [pos="0.283,0.038"]
Alcohol [pos="0.647,0.184"]
Anti_hypertensive_medication [pos="0.653,0.887"]
BMI [pos="0.214,0.778"]
CVD [outcome,pos="0.830,0.397"]
Cholesterol [pos="0.393,0.788"]
Glucose_lowering_medication [pos="0.801,0.746"]
HRV [exposure,pos="0.115,0.398"]
HbA1c [pos="0.231,0.635"]
Physical_activity [pos="0.595,0.019"]
Prior_CVD [pos="0.639,0.647"]
SES [pos="0.347,0.021"]
Sex [pos="0.232,0.073"]
Smoking [pos="0.676,0.090"]
Baseline_heart_rate [pos="0.114,0.205"]
Heart_rate [pos="0.432,0.325"]
Systolic_blood_pressure [pos="0.420,0.477"]
Age -> CVD
Age -> HRV
Alcohol -> CVD
Alcohol -> HRV
Anti_hypertensive_medication -> CVD
Anti_hypertensive_medication -> HRV
Anti_hypertensive_medication -> Systolic_blood_pressure
BMI -> CVD
BMI -> Cholesterol
BMI -> HRV
BMI -> HbA1c
Cholesterol -> CVD
Cholesterol -> HRV
Glucose_lowering_medication -> CVD
Glucose_lowering_medication -> HRV
HRV -> CVD
HRV -> Heart_rate
HRV -> Systolic_blood_pressure
HbA1c -> CVD
HbA1c -> Glucose_lowering_medication
HbA1c -> HRV
Physical_activity -> CVD
Physical_activity -> HRV
Prior_CVD -> Anti_hypertensive_medication
Prior_CVD -> CVD
Prior_CVD -> HRV
Prior_CVD -> Systolic_blood_pressure
SES -> CVD
SES -> HRV
Sex -> CVD
Sex -> HRV
Smoking -> CVD
Smoking -> HRV
Systolic_blood_pressure -> CVD
Baseline_heart_rate -> HRV
Heart_rate -> CVD

}
')

 plot( HRV_cycle_DAG )
```

```{r}
#| echo: false

HRV_cycle_DAG%>%
ggdag(text = FALSE) +
        geom_dag_node(color="darkgreen") +
    theme_dag() + theme(legend.position = "none") +
    geom_dag_label(color="black", size = 2.75) +
     scale_adjusted()   

```

```{r}
#| echo: false

dagitty('dag {
bb="0,0,1,1"
Age [pos="0.551,0.061"]
Alcohol [pos="0.257,0.164"]
Anti_hypertensive_medication [pos="0.785,0.715"]
BMI [pos="0.065,0.604"]
CVD [outcome,pos="0.830,0.397"]
Cholesterol [pos="0.197,0.756"]
Glucose_lowering_medication [pos="0.620,0.790"]
HRV [exposure,pos="0.306,0.407"]
HbA1c [pos="0.303,0.584"]
Physical_activity [pos="0.150,0.045"]
Prior_CVD [pos="0.637,0.593"]
SES [pos="0.683,0.057"]
Sex [pos="0.404,0.059"]
Smoking [pos="0.165,0.264"]
baseline_heart_rate [pos="0.173,0.404"]
heart_rate [pos="0.816,0.201"]
Systolic_blood_pressure [pos="0.368,0.730"]
Age -> CVD
Age -> HRV
Alcohol -> CVD
Alcohol -> HRV
Anti_hypertensive_medication -> CVD
BMI -> CVD
BMI -> Cholesterol
BMI -> HRV
BMI -> HbA1c
Cholesterol -> CVD
Cholesterol -> HRV
Glucose_lowering_medication -> CVD
HRV -> CVD
HRV -> Prior_CVD
HRV -> heart_rate
HRV -> Systolic_blood_pressure
HbA1c -> CVD
HbA1c -> Glucose_lowering_medication
HbA1c -> HRV
Physical_activity -> BMI
Physical_activity -> CVD
Physical_activity -> HRV
Prior_CVD -> Anti_hypertensive_medication
Prior_CVD -> CVD
SES -> CVD
SES -> HRV
Sex -> CVD
Sex -> HRV
Smoking -> CVD
Smoking -> HRV
Baseline_heart_rate -> HRV
heart_rate -> CVD
Systolic_blood_pressure -> Anti_hypertensive_medication
Systolic_blood_pressure -> CVD
Systolic_blood_pressure -> Prior_CVD
}') %>% 
ggdag(text = FALSE) +
        geom_dag_node(color="darkgreen") +
    theme_dag() + theme(legend.position = "none") +
    geom_dag_label(color="black", size = 2.75) +
     scale_adjusted()   

```

Minimal adjustments:

**Model 1: Age, Sex**

**Model 2: Age, SES, Sex, Alcohol, Physical_activity, Smoking, BMI, Cholesterol, HbA1c**

**Model 3: Age, SES, Sex, Alcohol, Physical_activity, Smoking, BMI, Cholesterol, HbA1c + Systolic_blood_pressure**

**Model 4: Age, SES, Sex, Alcohol, Physical_activity, Smoking, BMI, Cholesterol, HbA1c + Systolic_blood_pressure + Glucose_lowering_medication + Prior_CVD + Anti_hypertensive_medication**

In addition pre-adjusted HRV for baseline heart rate will be performed to correct for cealling effect

# HRV hour and CVD

```{r}
#| echo: false

DAG_HRV_hour <- dagitty('dag {
bb="0,0,1,1"
Age [pos="0.283,0.038"]
Alcohol [pos="0.674,0.092"]
Anti_hypertensive_medication [pos="0.625,0.847"]
BMI [pos="0.191,0.829"]
CVD [outcome,pos="0.788,0.426"]
Cholesterol [pos="0.401,0.817"]
Glucose_lowering_medication [pos="0.795,0.752"]
HRV_Hour [exposure,pos="0.141,0.408"]
HbA1c [pos="0.231,0.635"]
Heart_rate_Hour [pos="0.119,0.161"]
Physical_activity_Hour [pos="0.065,0.229"]
Prior_CVD [pos="0.639,0.647"]
SES [pos="0.347,0.021"]
Sex [pos="0.232,0.073"]
Smoking [pos="0.676,0.025"]
Systolic_blood_pressure [pos="0.441,0.472"]
Age -> CVD
Age -> HRV_Hour
Alcohol -> CVD
Alcohol -> HRV_Hour
Anti_hypertensive_medication -> CVD
Anti_hypertensive_medication -> HRV_Hour
BMI -> CVD
BMI -> Cholesterol
BMI -> HRV_Hour
BMI -> HbA1c
Cholesterol -> CVD
Cholesterol -> HRV_Hour
Glucose_lowering_medication -> CVD
HRV_Hour -> CVD
HRV_Hour -> Systolic_blood_pressure
HbA1c -> CVD
HbA1c -> Glucose_lowering_medication
HbA1c -> HRV_Hour
Heart_rate_Hour -> HRV_Hour
Physical_activity_Hour -> HRV_Hour
Prior_CVD -> Anti_hypertensive_medication
Prior_CVD -> CVD
Prior_CVD -> HRV_Hour
Prior_CVD -> Systolic_blood_pressure
SES -> CVD
SES -> HRV_Hour
Sex -> CVD
Sex -> HRV_Hour
Smoking -> CVD
Smoking -> HRV_Hour
Systolic_blood_pressure -> CVD
}'
)

plot(DAG_HRV_hour)
```

```{r}
#| echo: false

DAG_HRV_hour %>%
ggdag(text = FALSE) +
        geom_dag_node(color="darkgreen") +
    theme_dag() + theme(legend.position = "none") +
    geom_dag_label(color="black", size = 2.75) +
     scale_adjusted()  
```

```{r}
#| echo: false

dagitty('dag {
bb="0,0,1,1"
Age [pos="0.488,0.115"]
Alcohol [pos="0.304,0.068"]
Anti_hypertensive_medication [pos="0.693,0.633"]
BMI [pos="0.187,0.741"]
CVD [outcome,pos="0.788,0.426"]
Cholesterol [pos="0.235,0.530"]
Glucose_lowering_medication [pos="0.814,0.800"]
HRV_Hour [exposure,pos="0.408,0.417"]
HbA1c [pos="0.334,0.816"]
Heart_rate_Hour [pos="0.197,0.300"]
Physical_activity_Hour [pos="0.098,0.422"]
Prior_CVD [pos="0.502,0.514"]
SES [pos="0.715,0.127"]
Sex [pos="0.605,0.031"]
Smoking [pos="0.127,0.078"]
Systolic_blood_pressure [pos="0.408,0.663"]
Age -> CVD
Age -> HRV_Hour
Alcohol -> CVD
Alcohol -> HRV_Hour
Alcohol -> Heart_rate_Hour
Anti_hypertensive_medication -> CVD
BMI -> CVD
BMI -> Cholesterol
BMI -> HRV_Hour
BMI -> HbA1c
Cholesterol -> CVD
Cholesterol -> HRV_Hour
Glucose_lowering_medication -> CVD
HRV_Hour -> CVD
HRV_Hour -> Systolic_blood_pressure
HbA1c -> CVD
HbA1c -> Glucose_lowering_medication
HbA1c -> HRV_Hour
Heart_rate_Hour -> HRV_Hour
Physical_activity_Hour -> HRV_Hour
Physical_activity_Hour -> Heart_rate_Hour
Prior_CVD -> Anti_hypertensive_medication
Prior_CVD -> CVD
Prior_CVD -> HRV_Hour
SES -> CVD
SES -> HRV_Hour
Sex -> CVD
Sex -> HRV_Hour
Smoking -> CVD
Smoking -> HRV_Hour
Smoking -> Heart_rate_Hour
Systolic_blood_pressure -> Anti_hypertensive_medication
Systolic_blood_pressure -> CVD
}
'
) %>% 
    ggdag(text = FALSE) +
        geom_dag_node(color="darkgreen") +
    theme_dag() + theme(legend.position = "none") +
    geom_dag_label(color="black", size = 2.75) +
     scale_adjusted()  
```
