---
title: "analysis_backup"
format: html
editor: visual
---

```{r, include = F}
library(here)
source(here("R/library_packages.R"))
source(here("R/analysis.R"))
```

Individuals had the actiheart monitor for a least 48 hours

Include individuals with complete case

# Descriptive

```{r}
pro_data %>% 
    select(age_fup,sex,
               alchol_week_unit_total,PAEE_kj_kg_day,smoking_status,
               bmi,waist_av,fat_pc,
               p_lab_hba1c,
               p_lab_trig,p_lab_hdlc,p_lab_ldl,p_lab_chol,
               sbp_av,dbp_av
               ,SDNN, RMSSD) %>% 
    tbl_summary(
        statistic = list(
            all_continuous() ~ "{mean} ({sd})"
        ),
        missing = "no") %>% 
    as_flex_table()
```

## by sex

```{r}
pro_data %>% 
    select(age_fup,sex,
               alchol_week_unit_total,PAEE_kj_kg_day,smoking_status,
               bmi,waist_av,fat_pc,
               p_lab_hba1c,
               p_lab_trig,p_lab_hdlc,p_lab_ldl,p_lab_chol,
               sbp_av,dbp_av
               ,SDNN, RMSSD) %>% 
    tbl_summary(by = "sex",
        statistic = list(
            all_continuous() ~ "{mean} ({sd})"
        ),
        missing = "no") %>% 
    as_flex_table()
```

## Univariate analysis of SDNN

```{r}
# Univariate analysis

uni_var <- tbl_uvregression(
    pro_data[c("age_fup","p_gv_sex",
               "alchol_week_unit_total","PAEE_kj_kg_day","smoke_yn_num",
               "bmi","waist_av","fat_pc",
               "p_lab_hba1c",
               "p_lab_trig","p_lab_hdlc","p_lab_ldl","p_lab_chol",
               "sbp_av","dbp_av"
               ,"SDNN")
],
method = lm,
y = SDNN)

```

## Multivariate analysis of SDNN

```{r}
#Lifestyle markers
alchol_week_unit_total <- lm(SDNN ~ alchol_week_unit_total + p_gv_sex + age_fup , data= pro_data)
alchol_week_unit_total_est <- tbl_regression(alchol_week_unit_total, include = alchol_week_unit_total)

smoke_yn_num <- lm(SDNN ~ smoke_yn_num + p_gv_sex + age_fup , data= pro_data)
smoke_yn_num_est <- tbl_regression(smoke_yn_num, include = smoke_yn_num)

PAEE_kj <- lm(SDNN ~ PAEE_kj_kg_day + p_gv_sex + age_fup, data= pro_data)
PAEE_kj_est <- tbl_regression(PAEE_kj, include = PAEE_kj_kg_day)



# Obesity markers
waist_av <- lm(SDNN ~ waist_av + p_gv_sex + age_fup , data= pro_data)
waist_av_est <- tbl_regression(waist_av, include = waist_av)
fat_pc <- lm(SDNN ~ fat_pc + p_gv_sex + age_fup , data= pro_data)
fat_pc_est <- tbl_regression(fat_pc, include = fat_pc)
bmi <- lm(SDNN ~ bmi + p_gv_sex + age_fup, data= pro_data)
bmi_est <- tbl_regression(bmi, include = bmi)

# Glucose metabolism
p_lab_hba1c <- lm(SDNN ~ p_lab_hba1c + p_gv_sex + age_fup , data= pro_data)
p_lab_hba1c_est <- tbl_regression(p_lab_hba1c, include = p_lab_hba1c)

# Lipids
p_lab_trig <- lm(SDNN ~ p_lab_trig + p_gv_sex + age_fup , data= pro_data)
p_lab_trig_est <- tbl_regression(p_lab_trig, include = p_lab_trig)
p_lab_hdlc <- lm(SDNN ~ p_lab_hdlc + p_gv_sex + age_fup , data= pro_data)
p_lab_hdlc_est <- tbl_regression(p_lab_hdlc, include = p_lab_hdlc)

p_lab_ldl<- lm(SDNN ~ p_lab_ldl + p_gv_sex + age_fup, data= pro_data)
p_lab_ldl_est <- tbl_regression(p_lab_ldl, include = p_lab_ldl)

p_lab_chol<- lm(SDNN ~ p_lab_chol+ p_gv_sex + age_fup, data= pro_data)
p_lab_chol_est <- tbl_regression(p_lab_chol, include = p_lab_chol)

# Blood pressure

sbp_av <- lm(SDNN ~ sbp_av + p_gv_sex + age_fup, data= pro_data)
sbp_av_est <- tbl_regression(sbp_av, include = sbp_av)
dbp_av <- lm(SDNN ~ dbp_av + p_gv_sex + age_fup, data= pro_data)
dbp_av_est <- tbl_regression(dbp_av, include = dbp_av)



multi_var<- tbl_stack(list(alchol_week_unit_total_est,smoke_yn_num_est, 
                           PAEE_kj_est, 
                           bmi_est,waist_av_est,fat_pc_est,
                            p_lab_hba1c_est,
                           p_lab_trig_est ,p_lab_hdlc_est, p_lab_ldl_est, p_lab_chol_est,
                           sbp_av_est, dbp_av_est
                           ))
```

**Cardiovascular risk factors association with SDNN**

```{r}
tbl_merge(list(uni_var, multi_var),
          tab_spanner = c("Univariate", "Adjusted")) %>% 
    as_flex_table()
```

# Cardiovascular risk factors association with RMSSD

## Univariate analysis of RMSSD

```{r}
# Univariate analysis

uni_var_rmssd <- tbl_uvregression(
    pro_data[c("age_fup","p_gv_sex",
               "alchol_week_unit_total","PAEE_kj_kg_day","smoke_yn_num",
               "bmi","waist_av","fat_pc",
               "p_lab_hba1c",
               "p_lab_trig","p_lab_hdlc","p_lab_ldl","p_lab_chol",
               "sbp_av","dbp_av"
               ,"RMSSD")
],
method = lm,
y = RMSSD)

```

## Multivariate analysis of RMSSD

```{r}
# Obesity markers
waist_av <- lm(RMSSD ~ waist_av + p_gv_sex + age_fup , data= pro_data)
waist_av_est <- tbl_regression(waist_av, include = waist_av)
fat_pc <- lm(RMSSD ~ fat_pc + p_gv_sex + age_fup , data= pro_data)
fat_pc_est <- tbl_regression(fat_pc, include = fat_pc)
bmi <- lm(RMSSD ~ bmi + p_gv_sex + age_fup, data= pro_data)
bmi_est <- tbl_regression(bmi, include = bmi)

# Glucose metabolism
p_lab_hba1c <- lm(RMSSD ~ p_lab_hba1c + p_gv_sex + age_fup , data= pro_data)
p_lab_hba1c_est <- tbl_regression(p_lab_hba1c, include = p_lab_hba1c)

# Lipids
p_lab_trig <- lm(RMSSD ~ p_lab_trig + p_gv_sex + age_fup , data= pro_data)
p_lab_trig_est <- tbl_regression(p_lab_trig, include = p_lab_trig)
p_lab_hdlc <- lm(RMSSD ~ p_lab_hdlc + p_gv_sex + age_fup , data= pro_data)
p_lab_hdlc_est <- tbl_regression(p_lab_hdlc, include = p_lab_hdlc)

p_lab_ldl<- lm(RMSSD ~ p_lab_ldl + p_gv_sex + age_fup, data= pro_data)
p_lab_ldl_est <- tbl_regression(p_lab_ldl, include = p_lab_ldl)

p_lab_chol<- lm(RMSSD ~ p_lab_chol+ p_gv_sex + age_fup, data= pro_data)
p_lab_chol_est <- tbl_regression(p_lab_chol, include = p_lab_chol)

# Blood pressure

sbp_av <- lm(RMSSD ~ sbp_av + p_gv_sex + age_fup, data= pro_data)
sbp_av_est <- tbl_regression(sbp_av, include = sbp_av)

dbp_av <- lm(RMSSD ~ dbp_av + p_gv_sex + age_fup, data= pro_data)
dbp_av_est <- tbl_regression(dbp_av, include = dbp_av)



multi_var_rmssd <- tbl_stack(list(bmi_est, waist_av_est, fat_pc_est,
                           p_lab_hba1c_est,
                           p_lab_trig_est,p_lab_hdlc_est, p_lab_ldl_est, p_lab_chol_est,
                           sbp_av_est, dbp_av_est
                           ))

```

**Cardiovascular risk factors association with RMSSD**

```{r}
tbl_merge(list(uni_var_rmssd, multi_var_rmssd),
          tab_spanner = c("Univariate", "Adjusted")) %>% 
    as_flex_table()
```

**Cardiovascular risk factors association with SDNN and RMSSD**

```{r}
tbl_merge(list(uni_var,multi_var,uni_var_rmssd, multi_var_rmssd),
          tab_spanner = c("Univariate SDNN", "Adjusted SDNN","Univariate RMSSD", "Adjusted RMSSD")) %>% 
    as_flex_table()
```

**Lipids and glucose risk factors association with SDNN and RMSSD**

```{r}
# Glucose metabolism
p_lab_hba1c_m2 <- lm(SDNN ~ p_lab_hba1c + p_gv_sex + age_fup +  p_lab_trig +
                      p_lab_chol + alchol_week_unit_total + PAEE_kj_kg_day + smoke_yn_num + bmi, data = pro_data)

p_lab_hba1c_m2_est <- tbl_regression(p_lab_hba1c_m2)
p_lab_hba1c_m2_est

# Lipids
p_lab_trig_m2 <- lm(RMSSD ~ p_lab_trig + p_gv_sex + age_fup + alchol_week_unit_total +  PAEE_kj_kg_day + smoke_yn_num + bmi + p_lab_hba1c, data= pro_data)

p_lab_trig_m2_est <- tbl_regression(p_lab_trig_m2)
p_lab_trig_m2_est

p_lab_hdlc <- lm(RMSSD ~ p_lab_hdlc + p_gv_sex + age_fup , data= pro_data)
p_lab_hdlc_est <- tbl_regression(p_lab_hdlc, include = p_lab_hdlc)

p_lab_ldl<- lm(RMSSD ~ p_lab_ldl + p_gv_sex + age_fup, data= pro_data)
p_lab_ldl_est <- tbl_regression(p_lab_ldl, include = p_lab_ldl)

p_lab_chol<- lm(RMSSD ~ p_lab_chol + p_gv_sex + age_fup, data= pro_data)
p_lab_chol_est <- tbl_regression(p_lab_chol, include = p_lab_chol)
```
