---
title: "Actiheart_HRV_validation"
format: html
editor: visual
---

```{r}
#| include: false

library(dplyr)
library(purrr)
library(tidyr)
library(here)
library(vroom)
library(ggplot2)
source(here::here("R/hrv_functions/1_generate_IBI_data.R"))
source(here::here("R/hrv_functions/2_hrv_indices_calculations_functions.R"))
validation_HRV <- vroom(here::here("data/validation_HRV_8beat_data.csv"))

```


# Description of HRV calculation 

based on 30 seconds segment with mean heart rate and 95% prediction interval


## Data description

Interbeat intervals (IBI) from 24-hour heart rate recordings in 22 healthy indiviudals.
Further information about the data has been described elsewhere:

Rossi, A., Da Pozzo, E., Menicagli, D., Tremolanti, C., Priami, C., Sirbu, A., Clifton, D., Martini, C., & Morelli, D. (2020). Multilevel Monitoring of Activity and Sleep in Healthy People (version 1.0.0). PhysioNet. https://doi.org/10.13026/cerq-fc86.

## SDNN
```{r}
#| echo: false

validation_HRV %>%
ggplot(aes(gen_SDNN, SDNN)) +
    geom_point()

spearmann_sdnn <- cor.test(validation_HRV$gen_SDNN, validation_HRV$SDNN, method=c("spearman"))
spearmann_sdnn$estimate
```

## SDANN

```{r}
#| echo: false

validation_HRV %>%
    ggplot(aes(gen_SDANN, SDANN)) +
    geom_point()

spearmann_SDANN <- cor.test(validation_HRV$gen_SDANN, validation_HRV$SDANN, method=c("spearman"))
spearmann_SDANN$estimate
```

## SDNNi

```{r}
#| echo: false

validation_HRV %>%
    ggplot(aes(gen_SDNNIDX, SDNNIDX)) +
    geom_point()

spearmann_SDNNIDX <- cor.test(validation_HRV$gen_SDNNIDX, validation_HRV$SDNNIDX, method=c("spearman"))
spearmann_SDNNIDX$estimate
```

## RMSSD

```{r}
#| echo: false

validation_HRV %>%
    ggplot(aes(gen_RMSSD, RMSSD)) +
    geom_point()

spearmann_RMSSD <- cor.test(validation_HRV$gen_RMSSD, validation_HRV$RMSSD, method=c("spearman"))

spearmann_RMSSD$estimate
```

## HRVi

```{r}
#| echo: false

validation_HRV %>%
    ggplot(aes(gen_HRVi, HRVi)) +
    geom_point()

spearmann_HRVi <- cor.test(validation_HRV$gen_HRVi, validation_HRV$HRVi, method=c("spearman"))
spearmann_HRVi$estimate
```

## Mean HR

```{r}
#| echo: false

validation_HRV %>%
    ggplot(aes(gen_meanHR, meanHR)) +
    geom_point()
```


# Conclusion

HRV index based on distributed interbeat intervals were reliable and valid. HRV index based on succesive interbeat intervals showed more bias which make sence as we do not have or use order of heartbeat intervals.
