---
title: "Actiheart_HRV_validation"
format: html
editor: visual
---

```{r}
#| include: false

library(dplyr)
library(purrr)
library(tidyr)
library(here)
library(vroom)
library(ggplot2)
source(here::here("R/hrv_functions/1_generate_IBI_data.R"))
source(here::here("R/hrv_functions/2_hrv_indices_calculations_functions.R"))
validation_HRV <- vroom(here::here("data/validation_HRV_8beat_data.csv"))

```


# Description of HRV calculation 

based on 30 seconds segment with mean heart rate and 95% prediction interval


## Data description

from physionet

## SDNN
```{r}
#| echo: false

validation_HRV %>%
ggplot(aes(gen_SDNN, SDNN)) +
    geom_point()

spearmann_sdnn <- cor.test(validation_HRV$gen_SDNN, validation_HRV$SDNN, method=c("spearman"))
spearmann_sdnn$estimate
```

## SDANN

```{r}
#| echo: false

validation_HRV %>%
    ggplot(aes(gen_SDANN, SDANN)) +
    geom_point()

spearmann_SDANN <- cor.test(validation_HRV$gen_SDANN, validation_HRV$SDANN, method=c("spearman"))
spearmann_SDANN$estimate
```

## SDNNi

```{r}
#| echo: false

validation_HRV %>%
    ggplot(aes(gen_SDNNIDX, SDNNIDX)) +
    geom_point()

spearmann_SDNNIDX <- cor.test(validation_HRV$gen_SDNNIDX, validation_HRV$SDNNIDX, method=c("spearman"))
spearmann_SDNNIDX$estimate
```

## RMSSD

```{r}
#| echo: false

validation_HRV %>%
    ggplot(aes(gen_RMSSD, RMSSD)) +
    geom_point()

spearmann_RMSSD <- cor.test(validation_HRV$gen_RMSSD, validation_HRV$RMSSD, method=c("spearman"))

spearmann_RMSSD$estimate
```

## HRVi

```{r}
#| echo: false

validation_HRV %>%
    ggplot(aes(gen_HRVi, HRVi)) +
    geom_point()

spearmann_HRVi <- cor.test(validation_HRV$gen_HRVi, validation_HRV$HRVi, method=c("spearman"))
spearmann_HRVi$estimate
```

## Mean HR

```{r}
#| echo: false

validation_HRV %>%
    ggplot(aes(gen_meanHR, meanHR)) +
    geom_point()
```


# Conclusion

HRV index based on distributed interbeat intervals were valid. HRV index based on interbeat intervals showed more bias which make sence as we do not have or uss order of heartbeat intervals.
